---
output:
  pdf_document:
    latex_engine: xelatex
editor_options: 
  chunk_output_type: inline
---
```{r setup_5, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, results = 'asis')
```

Our aim is to gain a comprehensive understanding of the financial performance of
films on the basis of various factors. 
In this chapter, the investigation is based on a generalized additive model. 
We begin by examining the relationship between budget and revenue in different film genres, 
using both linear and non-linear modeling techniques. 
The number of variables considered that could have a possible influence is then 
expanded so that a generalized additive model can be created at the end. 
The model is analyzed for both scenarios, with and without interactions between the selected variables.

```{r library_5, include=FALSE}
# Step 1: Load all relevant libraries

library(readr) #
library(mgcv) #
library(dplyr) #
library(tidyr) #
library(MASS) #
library(ggplot2) #
library(scales) #
library(gridExtra)
library(kableExtra)
theme_set(theme_bw())
```

```{r data_imp, include=FALSE}
# We start by loading libraries and importing our dataset. 
# We then clean the data and perform transformations: 
  # extracting the first genre as the main gerne from the genre column
  # converting the financial variables to millions for better readability.

# Step 2: Import the dataset
TMBD_Movie_Dataset <- read_csv("../../00_Data/TMBD Movie Dataset.csv")

# Step 3: View the variables in detail & check the import
head(TMBD_Movie_Dataset)

# Check for missing values in the dataset
summary(TMBD_Movie_Dataset) # General NA and summary check
na_count <- colSums(is.na(TMBD_Movie_Dataset))
#print(na_count) # Print the number of NA values in each column
```

```{r data_prep, include=FALSE}
# Attention 1: Several factors for the categorical variable genres
#print(TMBD_Movie_Dataset$genres)

# Attention 2: Release date is separated with "-" between the values
#print(TMBD_Movie_Dataset$release_date)

# Step 3: Split the string "genres", focus only on the first/most relevant genre. 
# Therefore add a new column with the specific genre
# release_date handling as date and extract release year
movies <- TMBD_Movie_Dataset %>%
  mutate(release_year = as.integer(format(as.Date(release_date), "%Y")),
         First_Genre = sapply(strsplit(as.character(genres), "\\|"), `[`, 1))

# Set extracted Genre as factor
movies$First_Genre <- as.factor(movies$First_Genre)

# Step 4: set all values in millions in columns revenue, budget, profit
movies$revenue_adj <- movies$revenue_adj / 10^6
movies$budget_adj <- movies$budget_adj / 10^6
movies$profit <- movies$profit / 10^6

# check the dataset
# head(movies)
```

### Linear and Non-linear Models for Predicting Movie Revenue

As a first step, we start by fitting linear models for each genre to understand how well budget 
predicts revenue, evaluating the effectiveness of these models using \(R^2\) values.
The basic idea is, that the linear regression highlights the differences in 
predictive power between genres.

In a second step, the models were then adjusted to take into account the relationship 
between budget and revenue across all genres in order to minimise the influence of the number of observations.

In a third step we extend the analysis to non-linear relationships using
polynomial regression for quadratic effects, this analysis provides a deeper 
insight into the impact of budget on revenue with more complex patterns.
The following analysis carried out (entire section can be seen in the code) can 
be summarised with the following points:

```{r lm_analysis_5, fig.height = 6, fig.width = 10, fig.cap="R-squared values for linear models by genre", echo=FALSE, include = FALSE}
# Create an empty list to store all linear models for each genre
lm_results_1 <- list()
unique_genres <- unique(movies$First_Genre)

#log() for the predictor and response variable: They are amaounts and could be in a non-linear relationship
for (genre in unique_genres) {
  # Iterate with a for-loop over all movies with the same genre
  genre_data <- filter(movies, First_Genre == genre)
  
  # linear modell for the response variable "revenue-adj"  
  # based on the predictor "budget_adj"
  # meaning of "adj" = inflation adjustment into 2010 dollar values
  lm_model <- lm(log(revenue_adj) ~ log(budget_adj), data = genre_data)
  
  # Store the results inside an array
  lm_results_1[[as.character(genre)]] <- list(model = lm_model,
                                              summary = summary(lm_model))
}

# print all the r-squared values for each linear model
r_squared <- sapply(lm_results_1, function(x) x$summary$r.squared)
# Convert the R-squared values to a data frame for better printing
r_squared_df <- data.frame(Genre = names(r_squared), R_Squared = r_squared)
```

```{r plt_lm_5, fig.height=6, fig.width=10, fig.cap="Grapical analysis - linear model", echo=FALSE, include = FALSE}
# prepare the output for several plots
par(mfrow = c(3, 6))

# iterate over all models and plot each graph
for(genre in names(lm_results_1)) {
  # extract datas for each genre
  genre_data <- filter(movies, First_Genre == genre)
  
  # Plot
  plot(log(genre_data$budget_adj), log(genre_data$revenue_adj),
       main = paste(genre),
       xlab = "Budget", ylab = "Revenue",
       col = "blue", pch = 19)
  
  # add the regression line to the model: assumption for linearity
  abline(lm_results_1[[genre]]$model, col = "red")
}
```

```{r polynomial_model_5, echo = FALSE, include = FALSE }
# Create an empty list to store all polynomial models for each genre
poly_results <- list()

for (genre in unique_genres) {
  # Iterate with a for-loop over all movies with the same genre
  genre_data <- filter(movies, First_Genre == genre)
  
  # Polynomial model for the response variable "revenue_adj"
  # based on the predictor "budget_adj" with a quadratic term
  poly_model <- lm(log(revenue_adj) ~ log(budget_adj) + I(budget_adj^2), data = genre_data)
  
  #instead of I(...^2) also poly(..., degree = 2) is possible
  
  # Store the results inside an array
  poly_results[[as.character(genre)]] <- list(model = poly_model,
                                              summary = summary(poly_model))
}

# Extract the R-squared values for each polynomial model
r_squared_poly <- sapply(poly_results, function(x) x$summary$r.squared)

# Convert the R-squared values to a data frame for better printing
r_squared_poly_df <- data.frame(Genre = names(r_squared_poly), R_Squared = r_squared_poly)

```

```{r tbl_values_5, echo = FALSE, include = FALSE }
# Combine the linear and polynomial R-squared values into one data frame
combined_r_squared_df <- merge(r_squared_df, r_squared_poly_df, by = "Genre", suffixes = c("_Linear", "_Poly"))

# Print the combined table nicely using knitr::kable
knitr::kable(combined_r_squared_df, 
             caption = "$R^2$ values for linear and polynomial models by genre", 
             col.names = c("Genre", "$R^2$ (Linear)", "$R^2$ (Polynomial)"), 
             escape = FALSE) %>%
  kable_styling(position = "center", full_width = FALSE)


```


```{r graph_non_lm_5, fig.height=6, fig.width=10, fig.cap="Grapical analysis - non-linearity", echo=FALSE, include = FALSE}
# Create a list to store ggplot2 objects for nonlinear plots
gg_nonlinear_plots <- list()

# Iterate over all genres and create nonlinear plots
for(genre in unique_genres) {
  # Extract data for each genre
  genre_data <- filter(movies, First_Genre == genre)
  
  # Plotting the nonlinear relationship using ggplot2: assumption non-linearity
  gg_nonlinear_plots[[genre]] <- ggplot(data = genre_data, aes(x = log(budget_adj), y = log(revenue_adj))) +
    geom_point(color = "blue", size = 3) +
    geom_smooth(method = "lm", formula = y ~ poly(x, 2), color = "red") +
    labs(title = paste(genre), x = "Budget (in millions)", y = "Revenue (in millions)") +
    theme_minimal()
}

# Create a combined plot for all nonlinear plots
do.call(gridExtra::grid.arrange, c(gg_nonlinear_plots, ncol = 3))
```

**Genre-specific Analysis:**

  - The fit is highly dependent on which model (linear or non-linear) is used.
  - The \(R^2\) values vary significantly across genres, indicating that the budget's 
  explanatory power on revenue could be genre-dependent.
  - Some genres, like music and western show an almost perfect fit.
  - For other genres like horror and Adventure the \(R^2\) values are quite low.

This illustrates the strong effect of the number of observations, so that an 
overall model is better suited for general statements.

```{r graphical_analysis_5, fig.height=3, fig.width=8, fig.cap="Grapical analysis - linearity vs. non-linearity", echo=FALSE}
# Fit the linear model
lm.fit1 <- lm( log(revenue_adj) ~ log(budget_adj), data = movies)
#summary(lm.fit1)

# Fit the polynomial model
lm.fit2 <- lm(log(revenue_adj) ~ log(budget_adj) + I(budget_adj^2), data = movies)
#summary(lm.fit2)

# Plotting both models in one plot using ggplot2
ggplot(data = movies, aes(x = budget_adj, y = revenue_adj)) +
  geom_point(color = "black", size = 1) +
  geom_smooth(method = "lm", formula = y ~ x, color = "red", linewidth = 1, se = FALSE, linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), color = "seagreen", size = 1, se = FALSE, linetype = "solid") +
  labs(title = "Linear vs. Non-linear Relationship Model between Budget and Revenue", 
       x = "Budget [mill. USD]", y = "Revenue [mill. USD]") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values = c("Linear Model" = "red", "Polynomial Model" = "seagreen"))
```


**Linear Regression Model:**

  - The linear regression model indicates a positive relationship between the 
  budget and revenue of movies.
  - The coefficient for budget_adj is suggesting that for every million-dollar 
  increase in the budget, the revenue increases by approximately 0.9 million dollars.
  - Approximately 50% of the variance in revenue is explained by the budget alone and 
  the residuals exhibit a significant spread, indicating that many factors beyond budget influence movie revenue.

**Polynomial Regression Model (Quadratic Effect):**

  - This model has only a slightly higher \(R^2\) value.
  - Overall the model indicates that while there is a slight non-linear relationship, 
  the improvement over the linear model is only marginal.

**Graphical Analysis:**

  - The red dashed line represents the linear regression fit and many data points lie far away from the line.
  This is an indicator that the model does not capture all of the variability.
  - The green solid line represents the polynomial regression fit and captures some non-linear patterns, due to the slight bend.
  - However, the analysis has shown that the improvement is only marginally better than the linear model.

These findings suggest that further investigation of other potential influencing
factors and more complex modelling is necessary.
The next step will therefore be to examine the relationship between the running time 
of films and their revenues to see if there are any non-linear patterns that 
can be better captured by more flexible models.


### Is there an influence of the variable run-time?
The film duration is selected as a further influencing variable. The reason for 
selecting this variable is a recent survey[^3] which shows that society prefers 
films of a certain length, namely 92 minutes. Therefore, the film length should 
have a significant influence on the revenue and further increase the model quality.

Based on this assumption, the hypothesis is first examined using the t-test to 
determine whether the film length really has an influence on the revenue and can 
be included in the model.

1. **Null hypothesis (H0)**: There is no difference in revenue between films with 
a running time of more than 92 minutes and films with a running time of 92 minutes or less: \( H_0: \mu_1 = \mu_2 \)

```{r t-test_comments, echo = FALSE, include = FALSE}
# where \( \mu_1 \) is the average revenue of films with a running time of more than 
# 90 minutes and \( \mu_2 \) is the average revenue of films with a running time of 90 minutes or less.
```

2. **Alternative hypothesis (H1)**: Films with a running time of more than 90 minutes 
generate higher revenues than films with a running time of 90 minutes or less: \( H_1: \mu_1 > \mu_2 \)

```{r t-test runtime, echo = FALSE, include = FALSE}
# Erstellen von zwei Gruppen: mehr als 90 Minuten, 90 Minuten oder weniger
group1 <- subset(movies, runtime > 92)$revenue_adj
group2 <- subset(movies, runtime <= 92)$revenue_adj

# Durchführen des t-Tests
t.test_result <- t.test(group1, group2, alternative = "greater")  # 'greater' for a one sided test

# Print the t-test result as-is
print(t.test_result)
```
Here are the key results and their interpretation:

**t-value**: 5.06 & **p-value:** 3.096e-07

**Mean values of the samples:**

  - Mean value of group 1 (films > 90 minutes): 213
  - Mean value of group 2 (films < 90 minutes): 129

The t-value is positive and relatively high, so this could indicate a strong difference between the two groups.
The very small p-value (well below 0.05) means that the probability of observing 
such an extreme value of the test statistic if the null hypothesis were true is extremely low.
The mean values of each sample show that the average revenue of films with a running time of more 
than 90 minutes is around 84 millions higher than that of shorter films.

The results of the test support the alternative hypothesis that films with 
a running time of more than 92 minutes generate significantly higher revenues than 
films that last 92 minutes or less. The running time is therefore included in the model. 
As a next step, we create a multiple linear model with interactions with all variables mentioned above.

### Completion to a multiple linear model

The following differences are recognizable:
```{r multiple_lm, fig.height=4, fig.width=8, fig.cap="Multiple linear model", include=TRUE}
# Multiple linear model without interactions and without the genres
lm.fit_inter_ext_1 <- lm(log(revenue_adj) ~ log(budget_adj) * runtime, data = movies)
# summary(lm.fit_inter_ext_1)

# Multiple linear model with interactions and genres-specific
lm.fit_inter_ext_2 <- lm(log(revenue_adj) ~ log(budget_adj) * runtime * First_Genre, 
                         data = movies)
# summary(lm.fit_inter_ext_2)
```

```{r multiple_lm_plt, echo = FALSE, include = FALSE}
# Diagnostikplots
par(mfrow = c(1, 2))

plot(predict(lm.fit_inter_ext_1), resid(lm.fit_inter_ext_1))
abline(h = 0, col = "red")

plot(predict(lm.fit_inter_ext_2), resid(lm.fit_inter_ext_2))
abline(h = 0, col = "red")

plot(movies$revenue_adj, predict(lm.fit_inter_ext_1))
abline(0, 1, col = "red")

plot(movies$revenue_adj, predict(lm.fit_inter_ext_2))
abline(0, 1, col = "red")
```

  - Without genre-specific interactions, 50% of the variance in revenue is explained 
  and there are significant interactions between budget and duration, indicating 
  that these two variables together have an impact on revenue. 
  This model is simple and straightforward to describe.

  - With genre-specific interactions, the variance of the model is around 57% and 
  therefore higher than without interactions. There are some significant interactions between budget, 
  running time and genre, which indicates that these variables together have an influence on revenues.
  But this is depending on the genre and the model is far more complex.

These observations lead to the basic consideration that a more flexible model that 
can better capture non-linear relationships might be more appropriate. 
Therefore, the application of a Generalised Additive Model (GAM) appears to be a 
logical next step to model the relationships between variables more comprehensively 
and improve the accuracy of results. 

### GAM - with and without interactions 
The generalized additive model can be used to explore complex patterns that may 
not be fully captured by a linear modeling approach. Therefore, budget, genre, 
runtime and revenue are reflected in the model. The basic idea is to fit a GAM with a qualitative predictor, in this case Genre.
In a first model the approach is to observe the effect of budget, runtime and 
genre on revenue independently. Therefore a GAM with individual smooth terms for no interactions is builded.
 
```{r gam_1, fig.height=3, fig.width=8, fig.cap="GAM with Individual Smooth Terms (No Interactions)", echo = FALSE, include=TRUE}
# GAM with Individual Smooth Terms: no interactions
gam_model <- gam(log(revenue_adj) ~ First_Genre + s(budget_adj) + s(runtime), 
                 data = movies, family = gaussian())

par(mfrow = c(1, 2))
plot(gam_model, se = TRUE, col = "blue")
#summary(gam_model)

genre_counts <- table(movies$First_Genre)
#genre_counts
```

\newpage

For comparison, we extend the model and add another variable, the release year, to maximize the model, using interactions between the individual variables.

```{r gam2, fig.height=3, fig.width=8, fig.cap="GAM with Interaction Terms", echo = FALSE, include=TRUE }

#Notes for this analysis:
#Nun ein Modell für Interaktionen zwischen Budget und Genre 
par(mfrow = c(1, 2))
gam_model_interaktion <- gam(log(revenue_adj) ~ s(budget_adj) + s(runtime) + 
                               First_Genre + te(budget_adj, First_Genre, 
                               bs = c("tp", "re")) + te(release_year, 
                               First_Genre, bs = c("tp", "re")), 
                             data = movies, family = gaussian(), method = "REML")

plot(gam_model_interaktion)
#summary(gam_model_interaktion)
```


### Conclusion

The first approach of using a linear and multiple linear model to examine the 
dependency of the revenue has not proved to be very practicable. The influence of 
the factors is probably not linear and a model must be used that can reflect complex 
and non-linear relationships.The first Generalized Additive Model focuses on the individual effects of the budget
and genre on the adjusted revenue. Several genres show significant differences in 
revenue. The smooth term for budget is highly significant, indicating a non-linear 
relationship between budget and revenue. Overall the model explains about 50% of the variance in the revenue data.
The second model includes interaction terms, allowing it to capture more complex 
relationships between budget, genre, release year on the adjusted revenue.
All predictors show significant effects on revenue and the model explains 55% of the variance, 
which is a improvement.

The GAM models have shown that they are better able to capture the non-linear and 
interactive effects between different factors influencing movie revenues.
For the client, the results mean that future modeling efforts should continue 
to explore nonlinear and interactive effects to improve the accuracy and insight 
of predictions in the movie industry. By integrating additional variables, the explained 
variance could be increased to get a more accurate picture of which factors influence the financial success of movies.


[^3]: Talkerresearch.com (2024). Research reveals the perfect movie length. accessed on 14.05.2024 via  https://talkerresearch.com/research-reveals-the-perfect-movie-length/


