---
output:
  pdf_document: default
  fig_caption: yes
  html_document:
    df_print: paged
always_allow_html: true
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(forcats)
library(gridExtra)

theme_set(theme_bw())
```

```{r loading_datasets, include=FALSE}
data_raw <- read.csv("../../00_Data/TMBD Movie Dataset.csv", header = TRUE)
summary(data_raw)
```


The TMBD Movie Dataset contains information about `r nrow(data_raw)` different movies 
that were released between `r min(data_raw$release_year)` and `r max(data_raw$release_year)`. 
They are categorized in `r ncol(data_raw)` variables allowing an in-depth and representative analysis of the movie industry.
The data set can be accessed on kaggle and is publicly available[^1] .

## Fundamental insights
In the following, a further basic analysis of the data set is carried out, which 
is supported by various visual representations to give the client an initial insight into the data.

```{r num_movies_year, echo=FALSE}
# Time series plot: number of movies released each year
num_movies_year <- data_raw %>%
  group_by(release_year) %>%
  summarise(num_movies = n())

num_movies_plot <- ggplot(num_movies_year, aes(x = release_year, y = num_movies)) +
  geom_line(color = "darkgreen", linewidth = 1) +
  labs(x = "Release Year", y = "Number of Movies", title = "Number of movies released per year") +
  scale_x_continuous(breaks = seq(min(num_movies_year$release_year, na.rm = TRUE), max(num_movies_year$release_year, na.rm = TRUE), by = 10))

# key insights
total_movies <- sum(num_movies_year$num_movies)
most_movies_year <- num_movies_year %>% filter(num_movies == max(num_movies)) %>% pull(release_year)
most_movies_count <- max(num_movies_year$num_movies)

# Calculate year-over-year changes
num_movies_year <- num_movies_year %>%
  mutate(change = num_movies - lag(num_movies))

# Find the year with the greatest increase
max_growth_year <- num_movies_year %>%
  filter(change == max(change, na.rm = TRUE)) %>%
  pull(release_year)

max_growth_value <- max(num_movies_year$change, na.rm = TRUE)
```

```{r num_movies_year_plt, fig.height=3, fig.width=8, fig.cap= "Number of movies released", echo=FALSE}
num_movies_plot
```

Based on the data, there have been a total of `r total_movies` movies released over the years. 
The year with the most movie releases was `r most_movies_year` with a total of `r most_movies_count` movies released.
The year with the greatest increase in the number of movies released was `r max_growth_year`,
with an increase of `r max_growth_value` movies compared to the previous year.

```{r total_profit_year, echo=FALSE}
# Total Profit over Time
total_profit <- data_raw %>% 
  group_by(release_year) %>% 
  summarise(total_profit = sum(profit / 10^6, na.rm = TRUE))

plt.1_1 <- ggplot(total_profit, aes(x = release_year, y = total_profit)) +
  geom_line(color = "darkgreen", linewidth = 1) +
  labs(x = "Release Year", y = "Total Profit (in millions)", title = "Total Profit Over Time") +
  scale_y_continuous(labels = scales::label_number(scale = 1, accuracy = 0.1))
```


```{r movie_profit_year, echo=FALSE}
# Profit per movie over time
profit_by_year <- data_raw %>%  
  group_by(release_year) %>% 
  summarise(total_profit = sum(profit / 10^6, na.rm = TRUE), num_movies = n()) %>% 
  mutate(average_profit_per_movie = total_profit / num_movies)

plt.2_1 <- ggplot(profit_by_year, aes(x = release_year, y = average_profit_per_movie)) +
  geom_line(color = "darkgreen", linewidth = 1) +
  labs(x = "Release Year", y = "Average Profit (in millions)", title = "Profit Per Movie Over Time") +
  scale_y_continuous(labels = scales::label_number(scale = 1, accuracy = 0.1))
```

```{r movie_profit_plots,fig.height=3, fig.width=8, fig.cap="Profit over time ", echo=FALSE}
# Arrange the plots side by side
grid.arrange(plt.1_1, plt.2_1, ncol = 2)
```
```{r most_profitable_genre, include=FALSE}
genre_profit <- data_raw %>%
  separate_rows(genres, sep = "\\|") %>%
  group_by(genres) %>%
  summarise(mean_profit = mean(profit, na.rm = TRUE))

most_profitable_genre <- genre_profit %>%
  arrange(desc(mean_profit)) %>%
  slice(1) %>%
  pull(genres)
```
The average profit of the movies in the dataset amounts to $`r sprintf("%.2f", mean(data_raw$profit, na.rm = TRUE) / 1000000)` 
million, whereas the most profitable genre appears to be `r most_profitable_genre`.

```{r top10_revenue,fig.height=4, fig.width=8, fig.cap="Top performing movies", echo=FALSE}
# Sort the dataset by earnings in descending order
data_sorted <- data_raw[order(-data_raw$revenue), ]

# Select the top 10 movies
top_10_movies <- head(data_sorted, 10)

# Convert revenue to millions
top_10_movies$revenue_millions <- top_10_movies$revenue / 10^6

# Calculate the range of release years
min_release_year <- min(top_10_movies$release_year, na.rm = TRUE)
max_release_year <- max(top_10_movies$release_year, na.rm = TRUE)
median_release_year <- median(top_10_movies$release_year, na.rm = TRUE)

# Plot the top 10 movies
top_10_plot <- ggplot(top_10_movies, aes(x = reorder(original_title, revenue_millions), y = revenue_millions)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(x = "", y = "Revenue (in millions)", title = "Top 10 Movies by Revenue") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::label_number(scale = 1, accuracy = 0.1, suffix = "M"))

# Display the plot
top_10_plot

```

The top 10 movies by revenue were released between the years `r min_release_year` and `r max_release_year`, 
with a median release year of `r median_release_year`. It is clear to see that newer films generate higher revenues. 
Possibly higher budget values are available and these could have an influence on the revenue. 
The following chart also provides an interesting initial insight, showing the distribution of revenues per genre and the number of movies per genre

```{r count_by_genre, echo=FALSE}
data_clean_genre <- data_raw %>%
  separate_rows(genres, sep = "\\|")

# Frequency of each genre
genre_counts <- data_clean_genre %>%
  count(genres) %>%
  arrange(n)

# Reordering genres based on frequency
data_clean_genre$genres <- factor(data_clean_genre$genres, levels = genre_counts$genres)

plot1 <- ggplot(data_clean_genre, aes(x = genres)) + 
  geom_bar(fill = "skyblue") +
  labs(title = "Distribution of Movies by Genre", x= "", y ="Number of movies") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r top10_genre_ratings, echo=FALSE}
# Calculate average revenue for each genre
genre_revenue <- data_clean_genre %>%
  group_by(genres) %>%
  summarise(avg_revenue = mean(revenue / 10^6, na.rm = TRUE)) %>%
  arrange(desc(avg_revenue))

# Selecting top 10 genres by average revenue
top_10_genres_revenue <- head(genre_revenue, 10)

# Plotting top 10 genres by average revenue
plot2 <- ggplot(top_10_genres_revenue, aes(x = reorder(genres, avg_revenue), y = avg_revenue, fill = genres)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(x = "", y = "Average Revenue (in millions)", title = "Top 10 Genres by Average Revenue") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::label_number(scale = 1, accuracy = 0.1, suffix = "M"))


```

```{r genre_plts,fig.height= 3, fig.width=8, fig.cap="genre insights", echo=FALSE}
# Arrange the plots side by side
grid.arrange(plot1, plot2, ncol = 2)
```

It can be seen that the number of films per genre does not directly result in higher revenues.
For example, most films belong to the drama genre, but in terms of average revenues, 
the drama genre is not represented in the top 10. Based on the initial basic findings 
of the project groups and the requirements of the client, the defined hypotheses 
are examined using suitable models in the following chapters.

[^1]: https://www.kaggle.com/datasets/successikuku/tmbd-movie-dataset/data
